---
title: "team.Rmd"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
dat <- readxl::read_xls('GSS.xls')
print(dat)
```


```{r}
dat = rename(dat, mStatus = `Marital status`)
dat = rename(dat, highestSchool = `Highest year of school completed`)
dat = rename(dat, sex = `Respondents sex`)
dat = rename(dat, politicalParty = `Political party affiliation`)
dat = rename(dat, happiness = `General happiness`)
dat = rename(dat, age = `Age of respondent`)
```
```{r}
dat$highestSchool = as.numeric(as.character(dat$highestSchool))

```

1.


```{r}
temp <- dat %>%
  group_by(mStatus) %>%
  summarise(Ave = mean(highestSchool, na.rm = TRUE)) 

temp <- na.omit(temp)

ggplot(temp, aes(x=mStatus, weight=Ave)) + geom_bar()
```



<br />
2.
```{r}
highEd <- filter(dat, highestSchool >= temp$Ave & (mStatus == temp$mStatus))
str(highEd)
head(highEd)
```

<br />
3. 
```{r}
veryHappy <- filter(dat, happiness == 'Very happy')

ggplot(veryHappy, aes(x="", y=sum(happiness == 'Very happy'), fill=mStatus)) + geom_bar(stat="identity", width=1) + coord_polar("y", start=0)

```


```{r}
prettyHappy <- filter(dat, happiness == 'Pretty happy')

ggplot(prettyHappy, aes(x="", y=sum(happiness == 'Pretty happy'), fill=mStatus)) + geom_bar(stat="identity", width=1) + coord_polar("y", start=0)

```

```{r}
print(sum(prettyHappy$happiness == 'Pretty happy'))
print(sum(veryHappy$happiness == 'Very happy'))
```


As we can see from the pie charts, the large majority of the people that responded very happy are married.  This means that there is a greater chance that one is happier if they are married.  

<br />
4.
```{r}

ggplot(prettyHappy, aes(x="", y=sum(happiness == 'Pretty happy'), fill=mStatus)) + geom_bar(stat="identity", width=1) + facet_wrap(~ politicalParty)
```

<br />
5. 
```{r}
ggplot(dat, aes(x = politicalParty, y = highestSchool)) + geom_boxplot() + theme(axis.text.x = element_text(angle = 90))
```

<br />
6.
```{r}
ggplot(dat, aes(x=sex, fill=happiness))+geom_bar(position='fill')
```

Males and females reported very similar levels of happiness.

```{r}
ggplot(dat, aes(x=sex, y=highestSchool))+geom_boxplot()

```

Females and Males have similar mean for their highest level of education, but males have a higher third quartile than females.
